---
title: "EDS241: Assignment 1"
author: "Charles Hendrickson"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "openxlsx", "estimator")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation


```


```{r , include=TRUE}

# Load the CalEnviroScreen 4.0 data from the California Office of Environmental Health Hazards Assessment (OEHHA).

mydata <- read.xlsx("data/CES4.xlsx")

# Select the specific columns we will be using in our analysis. 

mydata <- mydata %>%
  select("Census.Tract", "Total.Population", "California.County", "Low.Birth.Weight", "PM2.5", "Poverty")

# omit all NA values from the dataset

mydata <- na.omit(mydata)

```

\noindent (a) What is the average concentration of PM2.5 across all census tracts in California?

```{r}
pm2.5_avg <- mydata %>%
  summarise(pm2.5_avg = mean(PM2.5))

print(pm2.5_avg)
```

\noindent (b) What county has the highest level of poverty in California?

```{r}
# Find the mean poverty level per county
county_poverty_means <- mydata %>%
  group_by(California.County) %>%
  summarise(mean_poverty = mean(Poverty))
  
# Find the maximum mean poverty level
max(county_poverty_means$mean_poverty)

# Filter for the county with the max mean poverty level of 51.45584
max_poverty <- county_poverty_means %>% 
  filter(mean_poverty >= 51.45584) %>% 
  summarise(California.County)

print(max_poverty)

```

\noindent (c) Make a histogram depicting the distribution of percent low birth weight and PM2.5.

```{r}
# Histogram for PM2.5 levels
ggplot(data = mydata, aes(x= PM2.5)) +
  geom_histogram() +
  labs(title = "Distribution of Annual Mean PM2.5 Levels for California Census Tract",
       x = "PM2.5 level",
       y = "Count") +
  theme_classic()
```

```{r}
# Histogram for percent low birth weight
ggplot(data = mydata, aes(x = Low.Birth.Weight)) +
  geom_histogram() +
  labs(title = "Distribution of Percent Low Birth Weight for California Census Tract",
       x = "Percent Low Birth Weight (%)",
       y = "Count") +
  theme_classic()
```

\noindent (d) Estimate a OLS regression of LowBirthWeight on PM25. Report the estimated slope coefficient and its heteroskedasticity-robust standard error. Interpret the estimated slope coefficient. Is the effect of PM25 on LowBirthWeight statistically significant at the 5%?

**The estimated slope coefficient is 0.118162. This means that for every one unit increase in PM2.5 levels, there is a 0.118162 percent increase in the percentage of low birth weights for the California census tract.**

**The standard error is 0.008334**

**The effect of PM25 on LowBirthWeight is statistically significant at the 5% level becuase our p-value is much lower than 0.05.**


```{r}
regression_results <- summary(lm(Low.Birth.Weight ~ PM2.5, data = mydata))

regression_results
```

















